Can historical data predict ecological responses to novel climate conditions?

### 2016-12-13 precip/manuscript/precip_ms.tex
more commits fun.
scaling the competition and size parameters for better fitting

\title{Weak interspecific interactions in a sagebrush steppe: Evidence from observational data, population models, and a removal experiment}

@@

\textbf{\large{Keywords:}} Coexistence, competition, integral projection model.

\bigskip \textbf{Running title:} Weak interspecific interactions

@@

\smallskip \textbf{Authorship statement:} ARK and PBA designed the experiment and supervised data collection. AT helped with the design of the data analysis.
\smallskip

@@

\begin{enumerate}
\item Climate is thought to be an important driver of population ecology however it can be difficult to determine exactly which aspects of climate affect populations and at what vital rates.  Tests of whether statistical models fit to climate and population data can accurately predict population responses to natural and experimental climate manipulations have rarely been conducted.   
\item Using long-term observational data collected from the 1920's to the 1950's from a sagebrush steppe plant community, previous research has reported a wide variety of climate effects on four dominant plant species. We tested whether statistical models fit to the historical data could predict the population response of plants in these original plots to natural and experimentally imposed climate variation some 60 years after the historical data collection ended. New plots in the same location where the historical data were collected were assigned to either drought or irrigation treatments imposed using rainfall shelters and automated sprinklers. The original plots were also monitored and served as a control. We modeled survival, growth and recruitment as functions of spring temperature and spring and summer soil moisture. We then used statistical models fitted to the historical data to predict the year and treatment effects observed in contemporary data.  
\item  For survival and recruitment, the treatment effects were never significantly positive, indicating that competitive release was not underestimated by the interspecific effects included in the baseline model. For  \textit{Poa secunda} recruitment, the removal treatment was actually negative, indicating that our baseline model overestimated competitive release.
\item For growth, the removal treatment effect was significant and positive for two species, \textit{Poa secunda} and \textit{Pseudoroegneria spicata}, indicating that the baseline model underestimated competitive release. However, including information about the location of these grass individuals with respect to removed \textit{Artemisia} failed to improve the growth regressions, raising questions about the mechanisms driving the positive response to removal.
\item For three of the four species, individual based models and integral projections models based on the vital rate regressions showed that removal treatment effects had little impact on population growth rates or simulated equilibrium cover following a species removal. For \textit{P. spicata}, the population models that included effects of \textit{Artemisia} removal projected greater competitive release than the baseline models. 
\item \emph{Synthesis}: Although models based on observational data did not perfectly predict species responses to removal, overall the results increase our confidence that competitive release is in fact small for at least three of our four study species. The implication is that interspecific resource competition is not the primary factor determining the dynamics of this community.

@@

We monitored soil moisture and air temperature in four of the precipitation experiment plot pairs using Decagon Devices (Pulman Washington) 5TM and EC-5 soil moisture sensors and 5TE temperature sensors.  We installed two soil moisture sensors in each monitored plot at 5 cm and two at 25 cm deep in the soil.  Air temperature was measured underneath the roofing of the shelter at 30 cm above ground. Data were logged automatically every four hours.

We conducted a simple statistical analysis to determine the effect of the rain-out shelters and irrigation on the observed soil moisture.

>>>>

\title{Can historical data predict ecological responses to novel climate conditions?}

@@

\textbf{\large{Keywords:}} Climate change, demographic models, rain-out shelter.

\bigskip \textbf{Running title:} Predicting climate response

@@

\smallskip \textbf{Authorship statement:} ARK and PBA designed the experiment and supervised data collection. AT helped with the data analysis.
\smallskip

@@

\begin{enumerate}
\item Climate is thought to be an important driver of population ecology however it can be difficult to determine exactly which aspects of climate affect populations and at which stages in an organisms life-cycle.  Moreover, there have been few tests of whether statistical models fit to historical climate and population data can accurately predict the future response of populations to natural and experimentally imposed climate conditions.   
\item Using long-term observational data collected from the 1920's to the 1950's from a sagebrush steppe plant community, previous research has reported a wide variety of climate effects on four dominant plant species. We tested whether statistical models fit to these historical observational data could predict the population response of these plant populations to natural and experimentally imposed climate variation some 60 years after the historical data collection ended. 
\\item We established sixteen new plots in the same location as the historical data and assigned each to either drought or irrigation treatments imposed using rainfall shelters and automatic sprinklers. The original plots were also monitored and served as a control. We fit statistical models for survival, growth and recruitment of each species as functions of seasonal temperature and soil moisture in the historical data.  We then used these models to predict the treatment effects observed in contemporary data.  
\item  For   \textit{Poa secunda} recruitment, the removal treatment was actually negative, indicating that our baseline model overestimated competitive release.
\item For growth, the removal treatment effect was significant and positive for two species, \textit{Poa secunda} and \textit{Pseudoroegneria spicata}, indicating that the baseline model underestimated competitive release. However, including information about the location of these grass individuals with respect to removed \textit{Artemisia} failed to improve the growth regressions, raising questions about the mechanisms driving the positive response to removal.
\item For three of the four species, individual based models and integral projections models based on the vital rate regressions showed that removal treatment effects had little impact on population growth rates or simulated equilibrium cover following a species removal. For \textit{P. spicata}, the population models that included effects of \textit{Artemisia} removal projected greater competitive release than the baseline models. 
\item \emph{Synthesis}: Although models based on observational data did not perfectly predict species responses to removal, overall the results increase our confidence that competitive release is in fact small for at least three of our four study species. The implication is that interspecific resource competition is not the primary factor determining the dynamics of this community.

@@

We monitored soil moisture and air temperature in four of the precipitation experiment plot pairs using Decagon Devices (Pulman Washington) 5TM and EC-5 soil moisture sensors and 5TE temperature sensors.  We installed two soil moisture sensors in each monitored plot at 5 cm and two at 25 cm deep in the soil.  Air temperature was measured underneath the roofing of the shelter at 30 cm above ground. We also installed sensors in a nearby area receiving ambient rainfall to serve as a measure of un-manipulated control conditions. Data were logged automatically every four hours.

We conducted a simple statistical analysis to determine the effect of the rain-out shelters and irrigation on the observed soil moisture. We averaged soil moisture readings by day plot group  and then fit a simple linear model to the soil moisture data to determine the effects
